---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Emily Peters
  - Abby Dunnigan
  - Meghan Catania
  - Teddy Packert
  - Collin Gates

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

About this project :clap:
Our group wanted to find a data set about a topic that we are all interested in. 

```{r}

athletes <- read.csv("C:/Users/peter/OneDrive/School 2023-24/Fall 23/DANL/Most_paid_athletes1.csv")

```


# Data

The data.frame `athletes` contains the top 10 highest payed athletes each year from 1990 to 2020 ranked by Forbes Magazine :trophy:


## Summary Statistics

```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(athletes) 
```


```{r}
skim(athletes) %>% 
  select(-n_missing)
```


## Mean Salary Per Year
This time series shows each mean `earnings....million.` by each `Year`.

```{r}
mean_salaries <- athletes %>% 
  group_by(Year) %>% 
  summarize(mean_sal = mean(earnings....million.))
```


```{r}
ggplot(data = mean_salaries) +
  geom_line(mapping=aes(x=Year, y=mean_sal))
   
```
It is clear from 1990 to 2020, the mean salary of the athletes increased significantly. This could account for an increasing demand in watching sports over time, or for inflation.

